---
title: "Hospital Data Report"
author: "Zach Christensen"
date: "September 4, 2015"
output: pdf_document
---

# Healthcare-Associated Infections, State 
The Healthcare-Associated Infections (HAI) measures - state data. These measures are developed by Centers for Disease Control and Prevention (CDC) and collected through the National Healthcare Safety Network (NHSN). They provide information on infections that occur while the patient is in the hospital. These infections can be related to devices, such as central lines and urinary catheters, or spread from patient to patient after contact with an infected person or surface. Many healthcare associated infections can be prevented when the hospitals use CDC-recommended infection control steps.

The metric reported is standardized infection ratios (SIRs). SIRs compare the actual number of Healthcare-Assiocated Infections at each hospital to the predicted number of infections. These ratios are adjusted for various risk factors and population traits. More information about SIRs and how they are calculated can be found here: http://www.leapfroggroup.org/media/file/SIRCalc.pdf.  .

## Import Libraries and Data Set
The analysis relies on several open source libraries, and uses data directly from the healthdata.gov API.
```{r, importData, warning=FALSE, message=FALSE}
# Load Required Libraries
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)

# Load data from healthdata.gov
states <- read.csv("http://data.medicare.gov/api/views/k2ze-bqvw/rows.csv?accessType=DOWNLOAD")
```


## Preprocessing and Exploration
The dataset containes `r ncol(states)`: `r colnames(states)`. Before the analysis, the data can be restructured and some redundant information removed.
```{r exploration, message=FALSE}
# Date Range
startDate <- unique(states$Measure.Start.Date)
endDate <- unique(states$Measure.End.Date)

# Split the Measure.ID column
states <- states %>% mutate("Measure" = substring(text = Measure.ID, first = 1, last = 5),
                            "Type" = substring(text = Measure.ID, first = 7, 
                                               last = length(Measure.ID)))

# List of different measures
measures <- unique(filter(states, Type == "SIR")[,c("Measure", "Measure.Name")])

# Then only select needed columns
states <- subset(states, select = c("State", "Measure.Name", "Measure", "Type", "Score"))
states <- dcast(states, State + Measure ~ Type)

# Join the measures so we get measure names with data
states <- left_join(states, measures)

# Remove rows which have NA
states <- states[-which(is.na(states$SIR)),]

# Create data.frame with just SIRs by state
sirs <- head(dcast(states, State ~ Measure, value.var = "SIR"))

# Preview Data
head(sirs[,1:5])
```

The data for this set was collected from `r startDate` to `r endDate`, with information about `r nrow(measures)` different measures collected from `r length(unique(states$State))` states or regions.The `r nrow(measures)` measures are various healthcare associated infections:
```{r echo=FALSE}
measures[,2]
```

## Analysis
We can plot the SIR scores for each Measure by state.
```{r, message=FALSE}
# Plot of Scores by state and Measure
qplot(State, SIR, data = states, col = Measure, main = "Plot of SIR scores by State and Measure") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position="bottom")
```

Visually, it looks like the the SIR scores for `r filter(measures, Measure == "HAI_2")` are usually the highest, while `r filter(measures, Measure == "HAI_1")` are the lowest. Below the average score is grouped by measure.
```{r, echo=FALSE}
states %>% group_by(Measure) %>% summarise(mean(SIR))
```

The visual 



```{r, correlation}
cor(sirs[,2:7])

```


# Confidence Intervals - values that contain 1
```{r}
nrow(filter(states, CI_LOWER < 1, CI_UPPER > 1))

```
